Robust receding horizon control for linear systems with state and input dependent disturbances

### 2015-04-22 draft_v2.tex
Preliminary synch.
Made some changes to separate the example cleanly.
Included explicit flop count.
sections 7-9

\usepackage{graphicx}      % include this line if your document contains figures
\usepackage{natbib}        % required for bibliography
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathrsfs}

@@

\begin{abstract}                % Abstract of not more than 250 words.
  A robust model predictive controller is proposed for constrained linear systems with unknown disturbances
  that are subject to state and input dependent piecewise affine bounds.
  % with a dependence on the system state and control input. 
  The formulation encompasses problems involving model uncertainty arising from unknown system
  parameters and linearisation errors. We propose a computationally efficient method for the online
  optimization of a receding horizon min-max cost over the class of feedback strategies subject robust
  constraint satisfaction and provide illustrative numerical examples.
\end{abstract}

@@

After describing the control problem in Section~\ref{sec:problem:statement} we summarise preliminary 
results which are necessary to formulate the considered problem explicitly (Section~\ref{sec:preliminaries}).
Section~\ref{sec:gen:mpQP} discusses the recursive solution for equality constrained min-max recursions, one of 
the essential parts of the active set solver, and this is applied to the min-max RMPC problem in Section~\ref{sec:applied:mpQP}
for the case that the active set of constraints is known. A line search to determine the active set of constraints is presented
in Section~\ref{sec:line:search}. The application of the proposed algorithm to an example problem is presented in 
Section~\ref{sec:example} and conclusions are drawn in Section~\ref{sec:conclusion}.

@@

@@

@@

\begin{figure}
\includegraphics[width=.48\textwidth]{res}
\caption{The light blue sets depict the projected state constraint iterates~$\mathcal X_m$,
    whereas the blue and pink lines are the state trajectories produced along the two line
    searches. The pink line search eventually violates the state constraints and terminates.}
\label{fig:numerical:example}
\end{figure}
\section{Conclusion}\label{sec:conclusion}

>>>>

\usepackage{graphicx}      % include this line if your document contains figures
\usepackage{natbib}        % required for bibliography
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{lpic}

@@

\begin{abstract}                % Abstract of not more than 250 words.
  A robust model predictive controller is proposed for constrained linear systems with unknown disturbances
  that are subject to state and input dependent piecewise affine bounds.
  % with a dependence on the system state and control input. 
  The formulation encompasses problems involving model uncertainty arising from unknown system
  parameters and linearisation errors. We propose a computationally efficient method for the online
  optimization of a receding horizon min-max cost over the class of feedback strategies subject robust
  constraint satisfaction and provide an illustrative numerical example.
\end{abstract}

@@

After describing the control problem in Section~\ref{sec:problem:statement} we summarise preliminary 
results which are necessary to formulate the considered problem explicitly in Section~\ref{sec:preliminaries}.
Section~\ref{sec:gen:mpQP} discusses the recursive solution for equality constrained min-max recursions, one of 
the essential parts of the active set solver, and this is applied to the min-max RMPC problem in Section~\ref{sec:applied:mpQP}
for the case that the active set of constraints is known. A line search to determine the active set of
constraints is described 
in Section~\ref{sec:line:search} and bounds on the computational complexity of the approach are discussed in
Section~7. An application of the proposed algorithm to an example problem is presented in 
Section~\ref{sec:example} and conclusions are drawn in Section~\ref{sec:conclusion}.

@@

\begin{subequations}\label{seq:sigma}
\begin{equation}\label{eq:sigma:normal}

@@

\begin{equation}\label{eq:sigma:degen}

@@

\section{Conclusion}\label{sec:conclusion}
We have proposed a receding horizon controller for linearly constrained linear systems with unknown
disturbances that are subject to state and input dependent bounds. The optimisation of predicted feedback laws
is performed using an efficient specialised active set solver. Future work will consider how to limit the number of active set changes while retaining
the guarantee of closed loop stability.

### 2015-04-22 example.tex
sections 7-9

%!TEX root = draft_v2.tex
\section{Example}\label{sec:example}
In this section we present a numerical simulation for a simplified levitating ball model, as derived in 
by~\cite{Schaich:2015}. The dynamics of the nonlinear system are given $m\ddot y = m g - c\frac{i^2}{y^2}$,
where $m$ denotes the mass of the ball, $g$ denotes the gravitation constant, $c$ denotes a constant depending
on the coil, $y$ the vertical distance to the coil, the simplification is due to neglecting inductive dynamics 
and assuming we that can influence the current~$i$ directly. Figure~\ref{fig:levitating:ball} illustrates the setup.
Choosing the state $x=(y,\dot y)$ and the input $u=i$ we obtain a state space representation which depends nonlinearly
on both the state and the input. An equilibrium is given for all $u=\sqrt{\frac{gm}{c}}x_1$.

Linearising the nonlinear differential 
equation $\dot x = f(x,u)$ around an equilibrium point $(\hat x, \hat
u)$ gives the approximate linear model 
%
\[
	 \Delta\dot{x} = \underbrace{\left(\begin{array}{cc}
	0 & 1 \\ \frac{2c\hat u^2}{m\hat x_1^3} & 0
	\end{array}\right)}_{\frac{\partial f}{\partial x}(\hat x,\hat
      u)}\Delta x 
+ \underbrace{\left(\begin{array}{c}
	0 \\ - \frac{2c\hat u}{m\hat x_1^2}
	\end{array}\right)}_{\frac{\partial f}{\partial u}(\hat x,\hat
      u)}\Delta u
\]
%
where $\Delta u = u -\hat{u}$ and $\Delta x \approx x-\hat{x}$.
We derive the discrete time dynamics with sampling rate $T_s$ using
the Euler formula $x^+=x+T_s f(x,u) =:\tilde f(x,u)$ giving
\[
\Delta x^+ = A \Delta x + B \Delta u , \quad
\Biggl\{\begin{aligned} A &= I+T_s\frac{\partial f}{\partial  x}(\hat
  x,\hat u) \\
B &= T_s \frac{\partial f}{\partial u}(\hat x,\hat u)
\end{aligned}
\]
%
A representation of the form~\eqref{eq:the:problem}

>>>>

%!TEX root = draft_v2.tex
\section{Complexity}\label{sec:complexity}
We next discuss bounds on the computational complexity of the algorithm outlined in Section~\ref{sec:line:search}.
%The computational complexity of the considered algorithm can easily be bounded. 
%In order to give implementation-free bounds  these are expressed 
%These are given in terms of floating point operations
%(flops). 
For a given active set at most $2N$ mpQPs must be solved, and this is done by decomposing the
problem as stated in section~\ref{sec:gen:mpQP}.  For each mpQP the main computational effort is to
determine $\Sigma$ in~\eqref{seq:sigma}. In the case of~\eqref{eq:sigma:normal} this requires a QR
decomposition with $2(mn^2-n^3/3)$ floating point operations (flops), where $m$ is the number of active constraints and $n$ the number of
optimisation variables, plus inversion of a triangular matrix with $(m^3-m)/6$ flops and several matrix
multiplications with $mpn$ flops (for the multiplication of a $m\times n$ matrix with a $n\times p$
matrix). The computationally more expensive case of~\eqref{eq:sigma:degen} requires a QR decomposition,
inversion of a definite matrix $W$ using Cholesky decomposition with $n^3/3$ flops, inversion of a triangular
matrix, several matrix multiplications and additions with $mn$ flops, and inversion of a definite matrix
$\Delta$. In the worst case this adds up to at most $11n^3+n^2-n/3$ flops to obtain $\Sigma$. The solution to
the problem is then the result of a matrix multiplication, and, in order to be able to solve the subsequent
mpQP, the matrix $W_\theta$ and the vector $c_\theta$ also need to be computed. Hence the overall flop count
in the worst case adds up to $11n^3+n^2+2/3n+2n^2 n_\theta$, where $n_\theta$ denotes the dimension of the
parameter~$\theta$. 
%At each stage two mpQPs have to be solved recursively, so that 
In the worst case the
recursion for the entire horizon therefore requires $N(2 n_x+5 n_x^2+101 n_x^3+8 n_x^2 (n_u+n_x))$ flops.

The number of active set changes needed to obtain the active set corresponding to the current plant state
is necessarily finite, since this number cannot exceed the total number of possible active sets. However upper
bounds on this number are unrealistically large (except in pathological cases) and hence of little use in bounding the number of times that the recursion of
Section~\ref{sec:line:search} needs to be solved. Instead we give the typical execution time for an example
system with an implementation in MATLAB  in section~\ref{sec:example}.

\section{Example}\label{sec:example}
In this section we present a numerical simulation for a simplified levitating ball model, as derived in~\cite{Schaich:2015}. The dynamics of the nonlinear system are given $m\ddot y = m g - c\frac{i^2}{y^2}$,
where $m$ denotes the mass of the ball, $g$ denotes the gravitation constant, $c$ denotes a constant depending
on the coil, $y$ the vertical distance to the coil, the simplification is due to neglecting inductive dynamics 
and assuming we that can influence the current~$i$ directly. Figure~\ref{fig:levitating:ball} illustrates the setup.
Choosing the state $x=(y,\dot y)$ and the input $u=i$ we obtain a state space representation which depends nonlinearly
on both the state and the input. For any $x_1>0$, the system is in equilibrium if $x_2=0$ and $u=\sqrt{\frac{gm}{c}}x_1$.

Linearising the nonlinear differential 
equation $\dot x = f(x,u)$ around an equilibrium point $(\hat x, \hat
u)$ gives the approximate linear model 
%
\[
	 \Delta\dot{x} = \underbrace{\left(\begin{array}{cc}
	0 & 1 \\ \frac{2c\hat u^2}{m\hat x_1^3} & 0
	\end{array}\right)}_{\frac{\partial f}{\partial x}(\hat x,\hat
      u)}\Delta x 
+ \underbrace{\left(\begin{array}{c}
	0 \\ - \frac{2c\hat u}{m\hat x_1^2}
	\end{array}\right)}_{\frac{\partial f}{\partial u}(\hat x,\hat
      u)}\Delta u
\]
%
where $\Delta u = u -\hat{u}$ and $\Delta x \approx x-\hat{x}$.
We derive the discrete time dynamics with sampling rate $T_s$ using
the Euler formula $x^+=x+T_s f(x,u) =:\tilde f(x,u)$ giving
\[
\Delta x^+ = A \Delta x + B \Delta u , \quad
\Biggl\{\begin{aligned} A &= I+T_s\frac{\partial f}{\partial  x}(\hat
  x,\hat u) \\
B &= T_s \frac{\partial f}{\partial u}(\hat x,\hat u)
\end{aligned}
\]
%
A representation of the form~\eqref{eq:the:problem}, in which the disturbance $w$ accounts for linearisation
errors, can be obtained by sampling input and state
constraints and using the mean-value theorem \citep[see][]{Schaich:2015}. This leads
to a disturbance representation of the form $\mathcal W(x,u)=\{w\in\mathbb R^n: 
\min_{k\in\mathcal S_w}\{ H_{k,i}^x x + H_{k,i}^u u\} \leq w_i \leq \max_{k\in\mathcal S_w} 
\{ H_{k,i}^x x + H_{k,i}^u u \} \}$,
where ${k\in\mathcal S_w}$ denotes the sample index and $i$ the row number. This point-wise polytopic representation
together with a polyhedral representation of state constraints allows us to compute a polyhedral
MRPI set~$\mathcal X^\infty$. As a terminal feedback controller $K$ we use the solution to the robust 
Lyapunov condition $V(x)-V((A+BK)x+Dw)\leq \gamma^2 w^Tw$ 
with $V(x)=x^T P_0 x\geq0$, i.e. $x^TP_0x - ((A+BK)x+Dw)^TP_0((A+BK)x+Dw)\geq x^T(Q+K^TRK)x -\gamma^2 w^T w$, 
\cite[see e.g.][]{Boyd:94}. This semi-definite program can also be used to find the minimum value of $\gamma^2$. Starting from 
a polyhedral MRPI set~$\mathcal X^\infty$ we apply the algorithm presented in section~\ref{sec:preliminaries}
to obtain the necessary stage constraints~$\mathcal Z_k$. At this stage we can introduce numerical values
for individual parameters: $T_s=30\,\text{ms}$, $C=1\,\frac{\text{kg m}^3}{\text{A}^2\text{s}^2}$,
$m=100\,\text{g}$, $\hat x_1 = 50\,\text{mm}$ and $\mathcal X=\{x:
\abs{x_1- \hat x_1}\leq 1\,\text{mm}\wedge \abs{x_2}\leq 105\,\frac{\text{mm}}{\text{s}}\}$, $\mathcal
U=\{u:\abs{ u-\hat u}\leq 10 \,\text{mA}\}$.
To obtain $\mathcal W(x,u)$ we sample $\mathcal X$ and $\mathcal U$ with a total of 25 samples, i.e. the cardinality
of $\mathcal S_w$ is 10. In the simulation we use horizon length $N=5$. We perform a cold started line search from $\chi_0=(0,0)$
to the outer bound of the constraint set at $\chi_e=(1,-30)$, the resulting trajectories at the boundaries of
individual active sets are plotted in figure~\ref{fig:plot}.

The simulation was implemented in MATLAB only using non-optimised built-in functions for the main computational
purposes. In the simulation illustrated in figure~\ref{fig:plot} there were a total of 10 iterations required,
out of which 7 correspond to changes of the dominating right hand side of $\mathcal W(x,u)$, for which no
recomputation of the solution is necessary. For $\chi_e$ on the boundary of the feasible set
the solution time is less than $0.5\, \text{s}$ in total (using a MacBook Pro with a $2.3\,\text{GHz}$ quad-core processor).
